### 2018-06-03 MSc_Marko_Docevski_tekst.tex
Updates from ShareLaTeX

@@

\\ GENJAM

@@

\cite{Sturm2016} All our LSTM networks have the same architecture, but operate over differ- ent vocabularies and are trained differently. One kind we build, which we term char-rnn, operates over a vocabulary of single characters, and is trained on a continuous text file. The second kind we build, folk-rnn, operates over a vocab- ulary of transcription tokens, and is trained on single complete transcriptions. We next discuss our training data, and then the architecture and training of our systems, and finally how we use them to generate new transcriptions.

@@

@@

x(1:T) be a sequence of tokens representing an encoded musical score, α ⊂ {1, 2, . . . ,T} a multi-index, and sup- pose ?
xα correspond to some fixed token values to be harmonized (e.g. a provided Soprano line). 

@@

@@

We propose two methods to control the trade-off between creativity and discipline of MidiNet. The first method is to manipulate the effect of the conditions by using them only in part of the intermediate transposed convolution layers of G, to give G more freedom from the imposed condi- tions. The second method capitalizes the effect of the fea- ture matching technique [27]: we can increase the values of λ1 and λ2 to make the generated music sounds closer to existing music (i.e. those observed in the training set).

@@

@@

>>>>

Првиот обид за алгоритамско креирање на музика е забележан уште во XVIII од страна на """Јохан Филип Кирнбергер - Секогаш спремен минует и поноњески композитор""" (гер. Der allezeit fertige Menuetten- und Polonaisencomponist)

@@

@@

\cite{Sturm2016} 
n particular, we build long short-term memory (LSTM) networks having
three hidden layers of 512 LSTM blocks each, and train them using approxi- mately 23,000 music transcriptions expressed with a textual vocabulary (ABC notation). We use this data because it is available, high-level with regards to the music it transcribes, and quite homogeneous with regards to the stylistic conventions of the music (it is crowd-sourced by musicians that play “session” music, e.g., Celtic, Morris, etc.).We take two approaches to training our models: one is character based, in which the system builds a model of joint probabilities of each textual character given the previous 50 characters; the other is “token” based, in which the system computes the joint probability of each token (which can be more than one character) given all previous tokens of a transcription. The result of training is a generative system that outputs transcriptions resembling those in the training material. Our practical aim is to create music transcription models that are useful in particular contexts of music composition, within and outside stylistic conventions particular to the training data. 
Our transcription data comes from a weekly repository of https://thesession. org/,8 an on-line platform for sharing and discussingmusic played in “traditional music sessions” (often Celtic and Morris).
An entry begins with two identifiers, followed by the title, tune type, meter, key, ABC code, date, and contributing user. Contributions vary in detail, with some being quite elaborate, e.g., specifying ornamentation, grace notes, slurs and chords. Most transcriptions are monophonic, but some do specify multiple voices. Many transcriptions have improper ABC formatting, are missing bar lines, have redundant accidentals, miscounted measures, and so on. We create data for training our char-rnn model in the following way.We keep
only five ABC fields (title, meter, key, unit note length, and transcription), and separate each contribution by a blank line. 
This leaves us with a text file having 13,515,723 characters in total, and 47,924 occurrences of T:.10 There are 135 unique characters, e.g., “A”, “:”, and “”, each of which becomes an element of the vocabulary for our char-rnn model.
We create data for training our folk-rnn model in the following way. We
remove title fields and ornaments. We remove all transcriptions that have fewer than 7 measures when considering repetitions (to remove contributions that are not complete transcriptions, but transcriptions of suggested endings, variations, etc.). We remove all transcriptions that have more than one meter or key.11 We transpose all remaining transcriptions (23,636) to a key with root C.
We impose a transcription token vocabulary — each token consists of one or more characters — for the following seven types (with examples in parens): meter (“M:3/4”), key (“K:Cmaj”), measure (“:|” and “|1”), pitch (“C” and “c’”),
grouping (“(3”), duration (“2” and “/2”), and transcription (“<s>” and “<\s>”). 
Our dataset has 4,056,459 tokens, of which 2,816,498 are pitch, 602,673 are du- ration, and 520,290 are measure. A majority of the 23,636 transcriptions consists of 150 tokens or fewer; and 75percent have no more than 190. There are 137 unique tokens, each of which becomes a vocabulary element for our folk-rnn model.

Each LSTM network we build has three hidden layers with 512 LSTM blocks each, and a number of input and output units equal to the number of characters or tokens in its vocabulary. We encode our transcriptions in a local fashion, like in [11, 36], where each element in the vocabulary is mapped to an input and output unit. (This is also called “one-hot encoding”.) The output of each network is a probability distribution over its vocabulary. The total number of parameters in our char-rnn model is 5,585,920; and that in our folk-rnn model is 5,621,722.
All our LSTM networks have the same architecture, but operate over differ- ent vocabularies and are trained differently. One kind we build, which we term char-rnn, operates over a vocabulary of single characters, and is trained on a continuous text file. The second kind we build, folk-rnn, operates over a vocab- ulary of transcription tokens, and is trained on single complete transcriptions. We next discuss our training data, and then the architecture and training of our systems, and finally how we use them to generate new transcriptions.
We initialise the learning rate to 0.002, and ap- ply a decay rate of 0.95 after the first 10 epochs.We build and train our folk-rnn model using our own implementation. This also employs the RMSprop algorithm, but with minibatches of 64 parsed transcriptions each. Since transcriptions in the dataset have different lengths (in number of tokens), we generate minibatches using a bucketing strategy, which places together in a minibatch sequences with approximately the same length, pads them to the maximum length using a “null” token, and then use a masking strategy to ignore null tokens when computing outputs and the loss function. We begin training with a learning rate of 0.003, and a rate decay of 0.97 applied after the 20 first epochs.
For both models, we clip gradients outside [−5, 5] to the limits, and employ
a dropout rate of 0.5 after each LSTM hidden layer. We train each model for 100 epochs in total. We use 95percent of the dataset as training data and 5percent as validation data (the latter for measuring progress in predicting characters or tokens). Through training, our char-rnn model learns a “language model” to produce ABC characters. On the contrary, our folk-rnn model learns a language model in a vocabulary more specific to transcription, i.e., a valid transcription begins with <s>, then a time signature token, a key token, and then a sequence of tokens from 4 types. Our folk-rnn model does not embody the ambiguity of meaning that char-rnn does, e.g., that C can mean a pitch, part of a pitch (C), a letter in a title (A Cup of Tea), or part of a key designation (K:Cmin).
With our trained models, it is a simple matter to have them generate output: we just sample from the probability distribution output by the model over its vocabulary, and use each selected vocabulary element as subsequent input. We can initialise the internal state of each model either randomly, or by inputing a valid “seed” sequence (e.g., beginning with <s>). Repeating the sampling process for N timesteps produces N characters/tokens in addition to the seed sequence.
The role of the composer is clear when we apply our system to create a new
piece of music in Sec. 4.3. Our intention behind seeding the system with the opening two bars of (Fig. 3) is to see how the system responds to an input that does not adhere closely to the stylistic conventions in its training data. Is it able to apply pattern variations even when the input pattern isn’t very close to the learned material? Through our experience, we find the knowledge embedded within the system translates into this different context with the guiding hand of the composer. Within this relatively restricted approach to composition, we find our systems useful for assisting in music material generation that goes in directions we thought little to take.

@@

\cite{Johnson2017} Product of Experts
Hinton (2002) proposed a system known as Product of Ex- perts (PoE) for combining multiple models of the same data. To combine a series of N experts with parame- ters Θ1,Θ2, . . . ,ΘN and associated probability distribu-
tions f1(·|Θ1), f2(·|Θ2), . . . , fN(·|ΘN), one can renormal- ize the product of their individual distributions.
Notice the similarity of this expres- sion to the definition of conditional probability: p(A|B) = p(A∩B) p(B) . In fact, this product can be interpreted as the con-
ditional probability that all experts choose x, given that all experts choose the same vector from the data space.
We use a pair of encodings, motivated
by the observation that good jazz melodies have both inter- esting contours as well as pitches that are sonorous with the chord progression. One encoding is based on intervals be- tween adjacent notes and the other is based on harmonic relationships with the current chord in the chord progres- sion. Each encoding is processed by a separate network component, and each component produces a candidate note probability distribution. These distributions are then com- bined using Product-of-Experts (Hinton, 2002) to produce a final distribution over notes that can be trained using the maximum-likelihood criterion.
The network is de- signed to receive the previous note played as input at each step, producing as output a set of probabilities for which note to play at the next time step. The network is trained using the maximum-likelihood objective (i.e. choose the network model that assigns the highest probability to the true training data).
In order to fully develop a LSTM-based model for jazz melodies, we needed to choose a representation for the musical data for the network to understand. Furthermore, dif- ferent representations might lead to different types of be- havior. For instance, if we simply represented the notes by a bit-vector where each bit corresponded to a single note, the model would not be able to generalize well to different in- tervals and other patterns, as it would have to learn relation- ships between each note separately. Franklin (2004) found that representations that were motivated by musical relation- ships led to better performance than more naive approaches.
To accomplish these goals, we decided to use two encod-
ings, one that focuses on intervals between successive notes, and the other that focuses on intervals of notes relative to the chord progression. We thus split our network into two network modules. The interval expert module receives an interval-based encoding, while the chord expert receives the chord-based encoding. This enables the network to learn particular relationships relative to the current trajectory of the melody (as learned by the interval expert) and also rela- tive to the current chord progression (as learned by the chord expert). Each of these modules produces a probability dis- tribution over which note to choose at the next time step. This probability distribution represents the level of belief that those notes should be chosen to play next. The encod- ings used are depicted in F2
Each expert receives as input:
1. the note chosen at the previous time step, encoded using the corresponding expert’s output format (so the interval expert receives the previous interval jump, and the chord expert receives the previous relative pitch class).
2. a beat vector giving the position of the current time step in a measure.
3. a position vector giving the position of the previous note relative to the upper and lower bounds.
4. a chord vector giving the notes of the current chord rela- tive to the expert’s current position.
Training
Our network is trained using the cross-entropy between the network predictions at each time step and the correct notes chosen in the training data. Equivalently, we want to maxi- mize (the log-likelihood of) the probability that the network outputs the training data perfectly, since we want the net- work to output improvisations that are similar to the training data. We can express the probability of a whole melodic seg- ment as a product of conditional distributions at each time step, i.e. the probability of generating a segment x from pa- rameters Θ can be factored as To train our network, we can then compute the gradient of the loss with respect to our parameters Θ by performing backpropagation through time (Werbos, 1990). In our experiments, each expert subnetwork was imple-
mented as two LSTM layers, each with 300 nodes. Dur- ing training, dropout of 0.5 was applied to the non-recurrent connections between layers (Srivastava et al., 2014; Pham et al., 2014). We used the ADAM optimizer, introduced by Kingma and Ba (2014), to automatically set the learning rate for our training procedure.
Generation
After our model is fully trained, we can use it to create new melodies from the learned probability distribution. At each time step, starting from the beginning of the piece, we com- pute the probability distribution output by our model. We then choose the next note to play proportionally to the prob- ability assigned to that note by our model. This note is then fed back into the model as the input for the next time step, and a probability distribution for the next time step is computed. This process repeats until we have created a full melody segment. Optionally, we can modify the probability distribution be-
fore sampling from it, in order to modify or constrain the output of the network. The first way to modify the distribu- tion is to vary the ratios between note probabilities.
Small values of x make the model more “adventurous” by causing the probabilities to become more similar to each other, and large values of x will cause the model to be “con- servative” by sampling high-probability notes more often and low-probability notes less often. Setting x = 1 results in the original probability distribution.

@@

x(1:T) be a sequence of tokens representing an encoded musical score, {1, 2, . . . ,T} a multi-index, and sup- pose ? correspond to some fixed token values to be harmonized (e.g. a provided Soprano line). 

@@

\cite{Mehri2016} Hence, our contribution is threefold:
1. We present a novel method that utilizes RNNs at different scales to model longer term de- pendencies in audio waveforms while training on short sequences which results in memory efficiency during training.
2. We extensively explore and compare variants of models achieving the above effect.
3. We study and empirically evaluate the impact of different components of our model on three audio datasets. Human evaluation also has been conducted to test these generative models.
SampleRNN helps to address this challenge by using a hierarchy of modules, each operating at a different temporal resolution. The lowest module processes individual samples, and each higher module operates on an increasingly longer timescale and a lower temporal resolution. Each module conditions the module below it, with the lowest module outputting sample-level predictions. The entire hierarchy is trained jointly end-to-end by backpropagation.
Rather than operating on individual samples, the higher-level modules in SampleRNN operate on non-overlapping frames of FS(k) (“Frame Size”) samples at the kth level up in the hierarchy at a time (frames denoted by f(k)). Each frame-level module is a deep RNN which summarizes the history of its inputs into a conditioning vector for the next module downward.
Because different modules operate at different temporal resolutions, we need to upsample each
vector c at the output of a module into a series of r(k) vectors (where r(k) is the ratio between the temporal resolutions of the modules) before feeding it into the input of the next module downward (Eq. 6). We do this with a set of r(k) separate linear projections.
Because different modules operate at different temporal resolutions, we need to upsample each
vector c at the output of a module into a series of r(k) vectors (where r(k) is the ratio between the temporal resolutions of the modules) before feeding it into the input of the next module downward (Eq. 6). We do this with a set of r(k) separate linear projections.
The lowest module (tier k = 1; Eqs. 7–9) in the SampleRNN hierarchy outputs a distribution over a sample xi+1, conditioned on the FS(1) preceding samples as well as a vector c(k=2)
from the i
next higher module which encodes information about the sequence prior to that frame. As FS(1) is usually a small value and correlations in nearby samples are easy to model by a simple memoryless module, we implement it with a multilayer perceptron (MLP) rather thanRNNwhich slightly speeds up the training. 
We use a Softmax because we found that better results were obtained by discretizing the audio signals (also see van den Oord et al. (2016)) and outputting a Multinoulli distribution rather than using a Gaussian or Gaussian mixture to represent the conditional density of the original real-valued signal. When processing an audio sequence, the MLP is convolved over the sequence, processing We use a Softmax because we found that better results were obtained by discretizing the audio signals (also see van den Oord et al. (2016)) and outputting a Multinoulli distribution rather than using a Gaussian or Gaussian mixture to represent the conditional density of the original real-valued signal. When processing an audio sequence, the MLP is convolved over the sequence, processing
We enable efficient training of our recurrent model using truncated backpropagation through time, splitting each sequence into short subsequences and propagating gradients only to the beginning of each subsequence. We experiment with different subsequence lengths and demonstrate that we are able to train our networks, which model very long-term dependencies, despite backpropagating through relatively short subsequences.
We implemented the WaveNet architecture as described in Oord et al. (2016). Ideally, we would have liked to replicate their model exactly but owing to missing details of architecture and hyper- parameters, as well as limited compute power at our disposal, we made our own design choices so that the model would fit on a single GPU while having a receptive field of around 250 milliseconds
Apart from reporting NLL, we conducted AB preference tests for random samples from four models trained on the Blizzard dataset. For unconditional generation of speech which at best sounds like mumbling, this type of test is the one which is more suited. Competing models were the RNN, SampleRNN (2-tier), SampleRNN (3-tier), and our implementation of WaveNet. The rest of the models were excluded as the quality of samples were definitely lower and also to keep the number of pair comparison tests manageable.

@@

We propose two methods to control the trade-off between creativity and discipline of MidiNet. The first method is to manipulate the effect of the conditions by using them only in part of the intermediate transposed convolution layers of G, to give G more freedom from the imposed conditions. The second method capitalizes the effect of the fea- ture matching technique [27]: we can increase the values of λ1 and 2 to make the generated music sounds closer to existing music (i.e. those observed in the training set).

@@

This work was inspired by the legendary article by Andrej Karpathy - The Unreasonable Effectiveness of Recurrent Neural Networks \cite{AndrejKarpathy2015}. In it Karpathy shows how a neural network consisting of 2 layers of 512 LSTM cells can generalize over several datasets of textual data: Paul Graham essay, Shakespeare plays, XML and Linux C/C++ source code. The resulting model can generate new sequences that are reasonably close to the source material and can almost fool a human, and in the case of C/C++ code some of the results compile. It is a character-level language model, i.e., it learns the texts letter by letter. Our first experiment was to adapt this model to the task of music generation.

Eck et al. in \cite{Eck2002} describe the first application of deep learning to music generation. Due to the limited computational power at the time they defined a very simple experiment, generalization over 12 bar blues in written form, with a maximum of 8 notes per bar. They represented music in slices of note-time, i.e., one unit of time is \(\frac{1}{8}\)-th note long. This representation was later named as piano roll. The more complex of their two experiments consisted of learning in parallel a chord progression and a melody line, allowing them to generate both. Their architecture consists of 8 blocks of 2 LSTM cells, of which 4 are dedicated to learning melody the other 4 to learning the chord progression. The chords sections have output connections that influence the melody section but not vice versa. This work is the inspiration for most of the following research in the filed, including our own.

As an expansion to the previous work in \cite{Eck2008} Eck et al. improve their model by using musically important temporal information from the pieces, allowing it to learn structure at different granularities. The said information is metrically significant and is provided to the network as time-delayed copies of the network inputs at intervals that depend on the piece's meter. For pieces where no meter information is provided, the authors developed a technique for algorithmic meter extraction. The architecture consists of parallel LSTM layer and standard feed-forward layer, which is supposed to speed-up local dependency discovery. We plan to investigate how and whether having metrical information embedded into the learning model impacts performance.

The architectures presented in \cite{Hadjeres2016} and \cite{Liang2017} are specifically designed to tackle Bach chorales. A hybrid architecture was presented in \cite{Hadjeres2016} consisting of 2 LSTM subnets, of which one learns the sequence in standard order and the other in reverse order and a feed-forward subnet that learns the sequence in standard order. The three subnets are combined in a single feed-forward network. There is a copy of the architecture for each voice in the chorales. The music is generated by pseudo-Gibbs sampling procedure. They treat the 4 voices separately, as monophonic music generation problems, and combine them into a polyphonic melody. In contrast, we aim to generate mixture of chords/single notes from a single network output. To augment the dataset they transposed all the chorales to all transpositions that fit in the tonal range present in the corpus. We intend to do similar transposition schemes. The authors of \cite{Liang2017} approached the same problem with a 3-layer LSTM network, to which they applied Grid-search hyper-parameter tuning extensively to find optimal network and parameter configurations. They used a very different data representation, encoding all the voices in a single sequence, ordered by descending pitch.

Boulanger-Lewandowski et al. in \cite{Boulanger-Lewandowski2014} attempted to combine the RNNs' ability to model sequences and Restricted Boltzmann Machines' (RBMs) ability to model probability distributions and apply them to music generation, in an architecture dubbed RNN-RBM. In this architecture the RBM part models note co-occurrence probability distributions at each time-step and the RNN part models the temporal sequences. We aim to achieve polyphony by modeling  the probability of each output of the network independently of one another, even though realistically they are not, and try to avoid using a hybrid solution like RNN-RBN, due to it being difficult to train and computationally expensive.

In \cite{Johnson2017}, another hybrid deep learning model, the LSTM layers try to learn the temporal dependencies in the music, followed by a set of feed forward layers that try to learn note co-occurrences (or the harmonic part). They used a piano roll very similar to our own, enriched with additional meta-data extracted from the pieces.The meta-data depends on the music being of very regular form, which our dataset is no, so we did not include any.

The work presented in \cite{Bretan2016} approaches music generation as a word-level language model, where each unit (composed of 1 to 4 bars of music) is treated as word. The words are embedded in a descriptor vector. A library of word (or units) is then created from the complete set of embeddings. A LSTM network learns sequences of the embeddings. Music is generated by first sequencing embeddings, which are then decoded to actual note sequences by selecting the most appropriate representative out of the library of units. We used character level embeddings in out second experiment.

The implementation presented in \cite{Sturm2015} is very close to the one of \cite{AndrejKarpathy2015}. The authors transformed a set of folk songs in ABC notation into a modified version of the ABC notation that's more suitable for deep learning, that makes it very easy to unambiguously distinguish the various types of tokens, dubbed folk-rnn. On top of that they built a character-level model RNN deep learning model. Those tokens were encoded in one-hot vectors fed to 3 layer LSTM network, which outputs probability distribution over the unique token set. 
The authors of \cite{Madjiheurem2016} used an Auto-encoder architecture to learn chord latent space embeddings. A sequence learner model generalizes over sequences of latent space embeddings. They presented a comparison of several types of chord representations. One of the compared representations is very similar to what we used in our experiments. We discuss more in section VI.

@@

\cite{Sturm2016} We create data for training our folk-rnn model in the following way. We
remove title fields and ornaments. We remove all transcriptions that have fewer than 7 measures when considering repetitions (to remove contributions that are not complete transcriptions, but transcriptions of suggested endings, variations, etc.). We remove all transcriptions that have more than one meter or key.11 We transpose all remaining transcriptions (23,636) to a key with root C.
We impose a transcription token vocabulary — each token consists of one or more characters — for the following seven types (with examples in parens): meter (“M:3/4”), key (“K:Cmaj”), measure (“:|” and “|1”), pitch (“C” and “c’”),
grouping (“(3”), duration (“2” and “/2”), and transcription (“<s>” and “<\s>”). 
Our dataset has 4,056,459 tokens, of which 2,816,498 are pitch, 602,673 are du- ration, and 520,290 are measure. A majority of the 23,636 transcriptions consists of 150 tokens or fewer; and 75percent have no more than 190. There are 137 unique tokens, each of which becomes a vocabulary element for our folk-rnn model.
